---
title: "Lab 6 - Multiple Regression"
author: "Jenni Putz"
date: "11/12/2021"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r, message=F}
library(pacman)
p_load(tidyverse, stargazer, broom)

wages <- read_csv("wages.csv")
```

Let's look at the data frame:
```{r}
head(wages,10)
```

## Multiple Regression in R
We know how to do simple OLS--doing multiple OLS is just one easy step further! Let's run a regression of log wages on education and experience:
```{r}
reg1 <- lm(lwage ~ educ + exper, data = wages)
summary(reg1)
```
R gives us the t-stat and p-value so we can easily conduct hypothesis tests. Suppose we want to test 
$$H_0\text{: } \beta_1 = 0$$ vs $$H_1\text{: } \beta_1 \neq 0$$ at the 5% level.
We can look at the p-value and see that $p < \alpha$ so we can reject the null hypothesis at 5%.

Like last lab, we can use stargazer to summarize our regression results. This time, let's keep the adjusted $R^2$ in our table.
```{r, warning = F}
stargazer(reg1, keep.stat = c("rsq", "adj.rsq", "n"), type = "text")
```

### Confidence Intervals
How do we get confidence intervals in R? Let's do a 95% confidence interval.

Option 1: Construct by hand. We have the standard error and we have $\hat{\beta}_1$. We need to get the critical value of t.
```{r}
# We can use the qt() function to get the critical value of t. First put in your significance level. 
# for a two-tail test, make sure to divide by 2
# then the degrees of freedom. we have 1604-3 degrees of freedom
qt(.05/2, 1601)
```
The critical value of t is 1.96. Now we can plug that in to our confidence interval formula:
$$ .080330 - 1.96(.005259) \leq \beta_1 \leq .080330 + 1.96(.005259) $$
Which simplifies to:
$$ .0700 \leq \beta_1 \leq .0906$$

Option 2: Let R tell us. The `tidy()` function will give us the confidence interval if we tell it!
```{r}
tidy(reg1, conf.int = T)
```
The confidence interval matches our by-hand solution! 

But this is for a 95%. How do we get `tidy()` to give us a different confidence level? What if we want a 90% confidence interval?
```{r}
tidy(reg1, conf.int = T, conf.level = .9)
```


## Omitted Variable Bias
```{r, warning=F}
reg1 <- lm(lwage ~ educ, data = wages)
reg2 <- lm(lwage ~ educ + exper, data = wages)
stargazer(reg1, reg2, keep.stat = c("rsq", "adj.rsq", "n"), type = "text")
```

What is the sign of the OVB? Use the formula:
$$ Bias = \beta_2 \frac{cov(X_1, X_2)}{var(X_1)} $$
$\beta_2 = .046$, which is positive. We can get the covariance by using the covariance function.
```{r}
cov(wages$educ, wages$exper)
```
There is a negative bias from omitting experience.

Let's do another regression.
```{r, warning = F}
reg3 <- lm(lwage ~ educ + exper + IQ, data = wages) # positive bias from omitting IQ (imperfect proxy for intelligence)
stargazer(reg1, reg2, reg3, keep.stat = c("rsq", "adj.rsq", "n"), type = "text")
```
How are IQ and education correlated? Check the OVB formula again.
Here, $\beta_3 = .004$ which is positive. We can check the covariance to get the sign of the bias:
```{r}
cov(wages$educ, wages$IQ)
```
IQ and education are positively correlated and there is a positive bias from omitting IQ.

We can keep adding variables to our regression:
```{r, warning = F}
reg4 <- lm(lwage ~ educ + exper + IQ + KWW, data = wages) 
reg5 <- lm(lwage ~ educ + exper + IQ + KWW + fatheduc, data = wages) 
reg6 <- lm(lwage ~ educ + exper + IQ + KWW + fatheduc + motheduc, data = wages) 
stargazer(reg1, reg2, reg3, reg4, reg5, reg6, keep.stat = c("rsq", "adj.rsq", "n"), type = "text")
```

Let's look a little closer at the models with father's education included. The coefficient is not statistically significant. Should we still include this variable in our regression?

**YES**
Remember, if we exclude variables that should be in the
model and they are correlated with the regressors that we
include, then we will have
omitted variable bias
-- Even if they aren’t correlated with the included variables, they
help us tighten up our standard errors
-- If theory or intuition says you should have a variable in the
model, then it should be in there!


## F-tests
Let's suppose we want to test to see if the coefficients on father's education and mother's education are equal to zero at the 5% level.
The null hypothesis is:
$$H_0\text{: } \beta_5 = \beta_6 = 0$$
The alternative hypothesis is:
$$H_1\text{: either } \beta_5 \neq 0 \text{ or } \beta_6 \neq 0$$

To do the test, we need to identify our *restricted* model and our *unrestricted* model. `reg4` is the restricted model and `reg6` is the unrestricted. 

Recall, the F-stat is equal to:
$$F = \frac{(R^2_u - R^2_r)/q}{(1-R^2_u)/(n-k-1)}$$

The easy way and have R do it for us: Load the `car` package to use this function:
```{r}
p_load(car)
linearHypothesis(reg6, c("fatheduc=0", "motheduc=0"))
```

